---
title: 'Ocean Health Index data'
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 2
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../ohiprep/src/templates/ohi_hdr.html'
pdf_document:
  toc: true
---

We provide most Ocean Health Index (OHI) data through Github, but some files are just too large.  These larger files are available from this location.  Specifically, the following data are available here:

* Spatial files describing the OHI regions
* Raster files of the pressure data (scaled from 0-1) used in the 2016 OHI global assessment
* Some supplementary files used in our analyses 

The raster files may be particularly useful for researchers conducting regional OHI assessments (or, other analyses) that require summarizing the data for different regions than we used.   

Resources are available for viewing, analzying, and manipulating spatial data in R from our [Github repository](https://github.com/eco-data-science/), specifically:

* [raster data](https://rawgit.com/eco-data-science/spatial-analysis-R/gh-pages/intro_spatial_data_R.html)
* [vector data](https://rawgit.com/eco-data-science/spatial_analysis2_R/gh-pages/spatial_analysis2_dec2015.html)



```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(grid)
library(gridExtra)
library(png)
library(DT)
library(knitr)

```

## OHI regions

###[Download the OHI global regions](https://ohi.nceas.ucsb.edu/data/data/regions.zip)!

This folder includes shapefiles in both the [Mollweide](https://en.wikipedia.org/wiki/Mollweide_projection) (regions_mol) and WGS 84 (regions_gcs) coordinate reference systems (Fig. 1). See this [primer](https://github.com/OHI-Science/ohi-science.github.io/raw/dev/assets/downloads/other/CRS.pdf) for more information about coordinate reference systems. These files can be opened in mapping programs such as ArcGIS, GRASS GIS, and R.  Table 1 includes information about the data fields included in these shapefiles.

```{r move regions, eval=FALSE, echo=FALSE}

###################################
# Move region files to ftp site
###################################

# OHI regions
source('../ohiprep/src/R/common.R')

#library(raster)
#library(rgdal)

## Mollweide and Geographic Coordinates shapefiles
regions_mol <- readOGR(dsn= file.path(dir_M, "git-annex/globalprep/spatial/d2014/data"), layer="regions_mol")
 
writeOGR(regions_mol, "../../shares/web/data/htdocs/data", "regions_mol", driver="ESRI Shapefile")

regions_gcs <- readOGR(dsn= file.path(dir_M, "git-annex/globalprep/spatial/d2014/data"), layer="regions_gcs")
 
writeOGR(regions_gcs, "../../shares/web/data/htdocs/data", "regions_gcs", driver="ESRI Shapefile")

png("figures/regions_gcs.png", width=400, height=300, bg="transparent")
plot(regions_gcs, lwd=0.5)
dev.off()

setwd("../../shares/web/data/htdocs/data") # NOTE: have to set working directory because you can't have paths in zipfile
zip(zipfile = "regions",
                                      files = c("regions_mol.dbf", "regions_gcs.dbf",
                                                 "regions_mol.prj", "regions_gcs.prj",
                                                 "regions_mol.shp", "regions_gcs.shp",
                                                "regions_mol.shx", "regions_gcs.shx"))

file.remove(c("regions_mol.dbf", "regions_gcs.dbf",
              "regions_mol.prj", "regions_gcs.prj",
              "regions_mol.shp", "regions_gcs.shp",
              "regions_mol.shx", "regions_gcs.shx"))

setwd("../../../../../frazier/ftp_data")

##########################################
## some other spatial files that might be useful
## for some people
###########################################

# raster
file.copy("../ohiprep/globalprep/spatial/d2014/data/rgn_mol_raster_1km/sp_mol_raster_1km.tif",
"../../shares/web/data/htdocs/data/sp_mol_raster_1km.tif")

# raster data
file.copy("../ohiprep/globalprep/spatial/d2014/data/rgn_mol_raster_1km/regionData.csv",
"../../shares/web/data/htdocs/data/regionData.csv")

# ocean
file.copy(file.path(dir_M, "git-annex/globalprep/spatial/ocean.tif"), "../../shares/web/data/htdocs/data/ocean.tif")

# 3nm
regions_mol_3nm <- readOGR(dsn= file.path(dir_M, "git-annex/globalprep/spatial/d2014/data"), layer="regions_offshore3nm_mol")
 
writeOGR(regions_mol_3nm, "../../shares/web/data/htdocs/data", "regions_offshore3nm_mol", driver="ESRI Shapefile")


# sea ice
ice <- raster(file.path(dir_M, "git-annex/Global/NCEAS-Pressures-Summaries_frazier2013/ice_mask_resampled"))
writeRaster(ice, "../../shares/web/data/htdocs/data/ice_mask_resampled.tif", format="GTiff", progress="text")

## put files in a zip
setwd("../../shares/web/data/htdocs/data")
zip(zipfile = "spatial_ohi_supplement",
               files = c("ocean.tif",
                         "regionData.csv",
                         "regions_offshore3nm_mol.dbf",
                         "regions_offshore3nm_mol.prj",
                         "regions_offshore3nm_mol.shp",
                         "regions_offshore3nm_mol.shx",
                         "sp_mol_raster_1km.tif",
                         "ice_mask_resampled.tif"))
file.remove(c("ocean.tif",
                         "regionData.csv",
                         "regions_offshore3nm_mol.dbf",
                         "regions_offshore3nm_mol.prj",
                         "regions_offshore3nm_mol.shp",
                         "regions_offshore3nm_mol.shx",
                         "sp_mol_raster_1km.tif",
                         "ice_mask_resampled.tif"))

setwd("../../../../../frazier/ftp_data")



```

**Figure 1. OHI regions**

For the Mollweide (A) and WGS 84 (B) coordinate reference systems.

```{r region figs, echo=FALSE, fig.width=10, fig.height=4}

mol <- rasterGrob(readPNG("figures/global_map_Index_2016_mol.png"), interpolate=TRUE)
mol_txt <- textGrob("A", gp=gpar(fontsize=12), just="left", x=0.15, y=0)

gcs <- rasterGrob(readPNG("figures/regions_gcs.png"), interpolate=TRUE)
gcs_txt <- textGrob("B", gp=gpar(fontsize=12), just="left", x=0.15, y=0)
# c <- textGrob("c", gp=gpar(fontsize=12), just="left", x=0.15, y=0)
# d <- textGrob("d", gp=gpar(fontsize=12), just="left", x=0.15, y=0)
grid.arrange(mol_txt, gcs_txt, mol, gcs, nrow=2, ncol=2, heights=c(.4, 4))

```


**Table 1. OHI region metadata**

The region shapefiles include the following fields.

Field     | Includes    | Description
------------ | ----------- | -----------------------------------
rgn_typ      |  eez (N = 239), eez-disputed (15), eez-inland (5), fao (15), land (229), land-disputed (8), land-noeez (40) | Only the "eez" and "fao" regions are included in the global assessments.  "fao" regions refer to the [FAO Major Fishing Areas](http://www.fao.org/fishery/area/search/en), excluding eez regions.  The "eez" regions are based on the [exclusive economic zone boundaries](http://www.marineregions.org) of 220 regions and the [CCAMLR areas](https://www.ccamlr.org/en/organisation/convention-area-%E2%80%93-technical-description) for 19 regions in Antarctica.
ant_type     | eez (N = 220), eez-ccamlr (19), eez-disputed (15), eez-inland (5), fao (15), land (220), land-ccamlr (9), land-disputed (8), land-noeez (40) | This is similar to the rgn_typ variable but designates the Antarctica regions ("eez-ccamlr")
rgn_id       | numeric region identifiers (1 - 337) | 255 = disputed regions; 260-277 = fao regions; 301-337 = landlocked regions; 213 = Antarctica CCAMLR regions
ant_id       | numeric region identifiers | Same as rgn_id, except that Antarctica CCAMLR regions are identified
rgn_nam      | e.g., Antarctica, Peru    | Name of region
rgn_key      | e.g., ATA, UMI            | Short identifying code for region
are_km2      | numeric (<1, 30604795)   | Area of region in km^2^

### Additional OHI region data

**[Download additional OHI region files](https://ohi.nceas.ucsb.edu/data/data/spatial_ohi_supplement.zip)**

This zipfile contains several spatial files that we use for global OHI assessments:

Filename   | Format   | Description
---------- | -------- | ----------------
sp_mol_raster_1km.tif | tif raster | ~ 1km resolution raster of OHI regions (generated using the Mollweide spatial file provided above)
regionData.csv | csv | links the cell ids in the sp_mol_raster_1km.tif to other region data
ocean.tif | tif raster | ocean mask used to prepare pressure files
ice_mask_resampled.tif | tif raster | mask used to exclude ice areas for some pressures
regions_offshore3nm_mol | shapefile | 3nm offshore area for OHI regions


## OHI pressure data
Some of the processed OHI pressure data are available at the raster scale, which can be downloaded from Table 2.  To obtain the pressure score for each OHI region we average the raster cells within each region.  The raster data can be extracted at different spatial scales, which may be useful for regional OHI assessments.  

**An [example R script](https://rawgit.com/OHI-Science/ftp_data/master/tutorial/extract_spatial_data.html) describes how to extract and summarize raster data for regions.**

*Resolution:* ~1km 

*Coordinate reference system:* Mollweide 

*Coverage:* Global

*Cell values:* Scaled from 0 (low) - 1 (high-pressure)

```{r pressure data, echo=FALSE, eval=FALSE}

#################################
## ocean acidification data
#################################

years <- 2011:2015

for(year in years){ # year= 2011
file.copy(file.path(dir_M, "git-annex/globalprep/prs_oa/v2016/output", paste0("oa_prs_layer_", year, ".tif")),
          file.path("../../shares/web/data/htdocs/data", paste0("oa_prs_layer_", year, ".tif")))
}


setwd("../../shares/web/data/htdocs/data")
zip(zipfile = "acid", files = c(paste0("oa_prs_layer_", years, ".tif")))
file.remove(c(paste0("oa_prs_layer_", years, ".tif")))
setwd("../../../../../frazier/ftp_data")

############################
## SST
############################

years <- c("2008-2012", "2007-2011", "2006-2010", "2005-2009", "2004-2008")

for(year in years){ # year= 2011
file.copy(file.path(dir_M, "git-annex/globalprep/prs_sst/v2016/output", paste0("sst_", year, "_1985-1989.tif")),
          file.path("../../shares/web/data/htdocs/data", paste0("sst_", year, "_1985-1989.tif")))
}

setwd("../../shares/web/data/htdocs/data") 
zip(zipfile = "sst", files = c(paste0("sst_", years, "_1985-1989.tif")))
file.remove(c(paste0("sst_", years, "_1985-1989.tif")))
setwd("../../../../../frazier/ftp_data")

###############################
## SLR
###############################

years <- c(2011:2015)

for(year in years){ # year= 2011
file.copy(file.path(dir_M, "git-annex/globalprep/prs_slr/v2016/output", paste0("slr_", year, ".tif")),
          file.path("../../shares/web/data/htdocs/data", paste0("slr_", year, ".tif")))
}

setwd("../../shares/web/data/htdocs/data") 
zip(zipfile = "slr", files = c(paste0("slr_", years, ".tif")))
file.remove(c(paste0("slr_", years, ".tif")))
setwd("../../../../../frazier/ftp_data")

###############################
## UV
###############################

years <- c("2007_2011", "2008_2012", "2009_2013", "2010_2014", "2011_2015")

for(year in years){ # year= 2011
file.copy(file.path(dir_M, "git-annex/globalprep/prs_uv/v2016/output", paste0("uv_", year, "-2005_2009.tif")),
          file.path("../../shares/web/data/htdocs/data", paste0("uv_", year, "-2005_2009.tif")))
}

setwd("../../shares/web/data/htdocs/data") 
zip(zipfile = "uv", files = c(paste0("uv_", years, "-2005_2009.tif")))
file.remove(c(paste0("uv_", years, "-2005_2009.tif")))
setwd("../../../../../frazier/ftp_data")

###############################
## Trash
###############################

file.copy(file.path(dir_M, "git-annex/globalprep/cw_pressure_trash/v2015/output/weight_rescale.tif"),
          file.path("../../shares/web/data/htdocs/data/trash_weight_rescale.tif"))


###############################
## Artisanal fishing
###############################

years <- c("2003-2007", "2004-2008", "2005-2009", "2006-2010")

for(year in years){ # year= 2011
file.copy(file.path(dir_M, "git-annex/globalprep/prs_fish/v2016/out/artisanal", paste0("art_press_", year, ".tif")),
          file.path("../../shares/web/data/htdocs/data", paste0("art_press_", year, ".tif")))
}

setwd("../../shares/web/data/htdocs/data") 
zip(zipfile = "art_lb", files = c(paste0("art_press_", years, ".tif")))
file.remove(c(paste0("art_press_", years, ".tif")))
setwd("../../../../../frazier/ftp_data")

###############################
## Commercial fishing high bycatch
###############################

years <- c("2003-2007", "2004-2008", "2005-2009", "2006-2010")

for(year in years){ # year= 2011
file.copy(file.path(dir_M, "git-annex/globalprep/prs_fish/v2016/out/high_bycatch", paste0("hb_fish_pressure_", year, ".tif")),
          file.path("../../shares/web/data/htdocs/data", paste0("hb_fish_pressure_", year, ".tif")))
}

setwd("../../shares/web/data/htdocs/data") 
zip(zipfile = "comm_fis_hb", files = c(paste0("hb_fish_pressure_", years, ".tif")))
file.remove(c(paste0("hb_fish_pressure_", years, ".tif")))
setwd("../../../../../frazier/ftp_data")

###############################
## Commercial fishing low bycatch
###############################

years <- c("2003-2007", "2004-2008", "2005-2009", "2006-2010")

for(year in years){ # year= 2011
file.copy(file.path(dir_M, "git-annex/globalprep/prs_fish/v2016/out/low_bycatch", paste0("lb_fish_pressure_", year, ".tif")),
          file.path("../../shares/web/data/htdocs/data", paste0("lb_fish_pressure_", year, ".tif")))
}

setwd("../../shares/web/data/htdocs/data") 
zip(zipfile = "comm_fis_lb", files = c(paste0("lb_fish_pressure_", years, ".tif")))
file.remove(c(paste0("lb_fish_pressure_", years, ".tif")))
setwd("../../../../../frazier/ftp_data")

###############################
## Chemical pollution
###############################

years <- c(2009:2012)

# this one is a loner
file.copy(file.path(dir_M, "git-annex/globalprep/prs_chem/v2016/int/chemical_pollution_2013_scaled.tif"),
          file.path("../../shares/web/data/htdocs/data/chemical_pollution_2013_scaled.tif"))

for(year in years){ # year= 2011
file.copy(file.path(dir_M, "marine_threats/impact_layers_2013_redo/impact_layers/work/land_based/before_2007/raw_global_results/Frazier", paste0("chemical_pollution_", year, "_scaled.tif")),
          file.path("../../shares/web/data/htdocs/data", paste0("chemical_pollution_", year, "_scaled.tif")))
}

years <- 2009:2013
setwd("../../shares/web/data/htdocs/data") 
zip(zipfile = "chem_pollution", files = c(paste0("chemical_pollution_", years, "_scaled.tif")))
file.remove(c(paste0("chemical_pollution_", years, "_scaled.tif")))
setwd("../../../../../frazier/ftp_data")

###############################
## Nutrient pollution
###############################

years <- c(2009:2012)

# this one is a loner
file.copy(file.path(dir_M, "git-annex/globalprep/prs_land-based_nutrient/v2016/int/global_plumes_fert_2013_log_scaled_extend.tif"),
          file.path("../../shares/web/data/htdocs/data/global_plumes_fert_2013_log_scaled_extend.tif"))

for(year in years){ # year= 2011
file.copy(file.path(dir_M, "marine_threats/impact_layers_2013_redo/impact_layers/work/land_based/before_2007/raw_global_results/Frazier", paste0("global_plumes_fert_", year, "_raw_log_extend_scaled.tif")),
          file.path("../../shares/web/data/htdocs/data", 
                    paste0("global_plumes_fert_", year, "_raw_log_extend_scaled.tif")))
}

setwd("../../shares/web/data/htdocs/data") 
zip(zipfile = "nutrient_pollution", files = c("global_plumes_fert_2013_log_scaled_extend.tif", paste0("global_plumes_fert_", years, "_raw_log_extend_scaled.tif")))
file.remove(paste0("global_plumes_fert_", years, "_raw_log_extend_scaled.tif"))
file.remove("global_plumes_fert_2013_log_scaled_extend.tif")
setwd("../../../../../frazier/ftp_data")


```

**Table 2. OHI pressure raster data**

Links to download processed raster data for several OHI pressures.  Explore the links included in the table to learn more about the original data sources and how the data were prepared.  The "name" field links to a longer description of the data. The "description" field includes links to data preparation files (with R code).  NOTE: these large files can take a while to download and require a lot of memory!    

```{r pressure datatable, echo=FALSE, warning=FALSE}

link <- "http://ohi-science.org/ohi-global/layers.html#"

pressures <- c("cc_acid", "cc_sst", "cc_slr", "cc_uv", "po_trash", "fp_art_lb",
                      "fp_com_hb", "fp_com_lb", "po_chemicals", "po_nutrients")

p_data <- read.csv("../ohi-global/eez_layers_meta_data/layers_eez_base.csv") %>%
  filter(layer %in% pressures) %>%
  mutate(p_link = paste0(link, layer)) %>%
  mutate(name = paste0("[", name, "](", p_link, ")"))

p_prep <- read.csv("../ohi-global/eez_layers_meta_data/layers_eez_methods.csv") %>%
  filter(layer %in% pressures) %>%
  filter(variable == "dataprep_url_2016") %>%
  group_by(layer) %>%
  mutate(number = 1) %>%
  mutate(number = cumsum(number)) %>%
  mutate(number = paste0("data prep ", number)) %>%
  mutate(number = gsub(" 1", "", number)) %>%
  mutate(method = paste0("[", number, "](", variable_data, ")")) %>%
  summarize(method = paste(method, collapse = ", "))

download_data <- data.frame(layer = c("cc_acid", "cc_sst", "cc_slr", "cc_uv", "po_trash", "fp_art_lb", 
                                      "fp_com_hb", "fp_com_lb", "po_chemicals", "po_nutrients"),
                            Download = c("[5 years: 2011-2015, 1.7GB](https://ohi.nceas.ucsb.edu/data/data/acid.zip)",
                                         "[5 time periods: 2004-2008 to 2008-2012, 9.3GB](https://ohi.nceas.ucsb.edu/data/data/sst.zip)",
                                         "[5 years: 2011-2015, 99MB](https://ohi.nceas.ucsb.edu/data/data/slr.zip)",
                                         "[5 time periods: 2007-2011 to 2011-2015,  1.5GB](https://ohi.nceas.ucsb.edu/data/data/uv.zip)",
                                         "[1 year: 2015, 1.2GB](https://ohi.nceas.ucsb.edu/data/data/trash_weight_rescale.tif)",
                                         "[4 time periods: 2003-2007 to 2006-2010,  1.5GB](https://ohi.nceas.ucsb.edu/data/data/art_lb.zip)",
                                         "[4 time periods: 2003-2007 to 2006-2010,  2.6GB](https://ohi.nceas.ucsb.edu/data/data/art_lb.zip)",
                                         "[4 time periods: 2003-2007 to 2006-2010,  2.7GB](https://ohi.nceas.ucsb.edu/data/data/art_lb.zip)",
                                         "[5 years: 2009-2013, 2.3GB](https://ohi.nceas.ucsb.edu/data/data/chem_pollution.zip)",
                                         "[5 years: 2009-2013, 566MB](https://ohi.nceas.ucsb.edu/data/data/nutrient_pollution.zip)"))

p_data <- p_data %>%
  left_join(p_prep, by="layer") %>%
  mutate(description = paste0(description, " (", method, ")")) %>%
  left_join(download_data, by="layer") %>% 
  dplyr::select(Name=name, Description=description, Download)

kable(p_data)

```


## Mangrove data

###[Download our modified mangrove data](https://ohi.nceas.ucsb.edu/data/data/mangrove_2012_mw_km2.tif)

This is a slightly modified version of the ["MFW, Mangrove Forest Cover Loss since 2000 in areas with year 2000 Mangrove Forest"](http://faculty.salisbury.edu/~sehamilton/mangroves/index.html) dataset from:

Hamilton, S. E., & Casey, D. (2016). Creation of a high spatio-temporal resolution global database of continuous mangrove forest cover for the 21st century (CGMFC-21). Global Ecology and Biogeography, 25(6), 729-738. doi:10.1111/geb.1244.

Because the 30m resolution raster the data is provided as (although awesome!) can be difficult to work with, we created a ~500m resolution raster and converted to a Mollweide coordinate reference system.  The units are km^2^ mangrove per cell, consequently, summing the raster cells in a region will provide the km^2^ area of mangrove forest (cell values range from 0 to 0.274979).  

The scripts we used to estimate mangrove extent and trend are available from [github](https://github.com/OHI-Science/ohiprep/tree/master/globalprep/hab_mangrove/v2015).

These data could be useful for estimating mangrove cover in regions without local mangrove surveys.  

```{r mangroves, echo=FALSE, warning=FALSE, eval=FALSE}

file.copy(file.path(dir_M, "git-annex/globalprep/Mangrove/v2015/tmp/mangrove_2012_mw/mangrove_2012_mw_km2.tif"),
          "../../shares/web/data/htdocs/data/mangrove_2012_mw_km2.tif")

```

## Net Primary Productivity data

###[Download our modified NPP data](https://ohi.nceas.ucsb.edu/data/data/npp.zip)

We use Net Primary Productivity data to scale the fishing pressures data for the OHI global assessment.  

This is a slightly modified version of the [data](http://www.science.oregonstate.edu/ocean.productivity/standard.product.php) provided by:

Behrenfeld, M.J. and Falkowski, P.G., 1997. Photosynthetic rates derived from satellite‐based chlorophyll concentration. Limnology and oceanography, 42(1), pp.1-20.

As described [here](https://cdn.rawgit.com/OHI-Science/ohiprep/master/globalprep/prs_fish/v2016/prim_productivity/npp.html), we convert these data to a tif raster in the Mollweide coordinate reference system and average the monthly data to estimate mean annual NPP. Units are mg C/m2/data.  Data are available from 2003 to 2015.

These data could be useful for estimating fishing pressures at different spatial scales.  

```{r npp, echo=FALSE, warning=FALSE, eval=FALSE}

library(stringr)

npp_files <- list.files(file.path(dir_M, 'git-annex/globalprep/prs_fish/v2016/VGPM_primary_productivity/int/annual_npp'), pattern = 'annual_mean_npp_2', full.names = T)
years <- str_sub(npp_files, start=-8, end=-5) 

for(npp in npp_files){ #npp=npp_files[1]

year <-  str_sub(npp, start=-8, end=-5) 
  
file.copy(npp,
          sprintf("../../shares/web/data/htdocs/data/annual_mean_npp_%s.tif", year))
  }

setwd("../../shares/web/data/htdocs/data") 
zip(zipfile = "npp", files = c(paste0("annual_mean_npp_", years, ".tif")))
file.remove(paste0("annual_mean_npp_", years, ".tif"))
setwd("../../../../../frazier/ftp_data")

```


