---
title: 'Ocean Health Index data'
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 2
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../ohiprep/src/templates/ohi_hdr.html'
pdf_document:
  toc: true
---


This site provides links to the following Ocean Health Index data:

* Regions used in OHI global assessments
* Pressure raster data used in the 2016 OHI global assessment

Resources are available for viewing, analzying, and manipulating spatial data in R from our [Github repository](https://github.com/eco-data-science/), specifically:

* [raster data](https://rawgit.com/eco-data-science/spatial-analysis-R/gh-pages/intro_spatial_data_R.html)
* [vector data](https://rawgit.com/eco-data-science/spatial_analysis2_R/gh-pages/spatial_analysis2_dec2015.html)

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(grid)
library(gridExtra)
library(png)
library(DT)
library(knitr)

```

## OHI regions

[Download the OHI global regions](https://mazu.nceas.ucsb.edu/pressures/data/regions.zip)!

This folder includes shapefiles in both the [Mollweide](https://en.wikipedia.org/wiki/Mollweide_projection) (regions_mol) and latitude/longitude (regions_gcs) coordinate reference system (Fig. 1). See this [primer](https://github.com/OHI-Science/ohi-science.github.io/raw/dev/assets/downloads/other/CRS.pdf) for more information about coordinate reference systems. These files can be opened in mapping programs such as ArcGIS, GRASS GIS, and R.  Table 1 includes information about the data fields included in the shapefiles.

```{r move regions, eval=FALSE, echo=FALSE}

###################################
# Move region files to ftp site
###################################

# OHI regions
source('../ohiprep/src/R/common.R')

#library(raster)
#library(rgdal)

## Mollweide and Geographic Coordinates shapefiles
regions_mol <- readOGR(dsn= file.path(dir_M, "git-annex/globalprep/spatial/d2014/data"), layer="regions_mol")
 
writeOGR(regions_mol, "../../shares/web/pressures/htdocs/data", "regions_mol", driver="ESRI Shapefile")

regions_gcs <- readOGR(dsn= file.path(dir_M, "git-annex/globalprep/spatial/d2014/data"), layer="regions_gcs")
 
writeOGR(regions_gcs, "../../shares/web/pressures/htdocs/data", "regions_gcs", driver="ESRI Shapefile")

png("figures/regions_gcs.png", width=400, height=300, bg="transparent")
plot(regions_gcs, lwd=0.5)
dev.off()

setwd("../../shares/web/pressures/htdocs/data") # NOTE: have to set working directory because you can't have paths in zipfile
zip(zipfile = "regions",
                                      files = c("regions_mol.dbf", "regions_gcs.dbf",
                                                 "regions_mol.prj", "regions_gcs.prj",
                                                 "regions_mol.shp", "regions_gcs.shp",
                                                "regions_mol.shx", "regions_gcs.shx"))

file.remove(c("regions_mol.dbf", "regions_gcs.dbf",
              "regions_mol.prj", "regions_gcs.prj",
              "regions_mol.shp", "regions_gcs.shp",
              "regions_mol.shx", "regions_gcs.shx"))

setwd("../../../../../frazier/ftp_data")

##########################################
## some other spatial files that might be useful
## for some people
###########################################

# raster
file.copy("../ohiprep/globalprep/spatial/d2014/data/rgn_mol_raster_1km/sp_mol_raster_1km.tif",
"../../shares/web/pressures/htdocs/data/sp_mol_raster_1km.tif")

# raster data
file.copy("../ohiprep/globalprep/spatial/d2014/data/rgn_mol_raster_1km/regionData.csv",
"../../shares/web/pressures/htdocs/data/regionData.csv")

# ocean
file.copy(file.path(dir_M, "git-annex/globalprep/spatial/ocean.tif"), "../../shares/web/pressures/htdocs/data/ocean.tif")

# 3nm
regions_mol_3nm <- readOGR(dsn= file.path(dir_M, "git-annex/globalprep/spatial/d2014/data"), layer="regions_offshore3nm_mol")
 
writeOGR(regions_mol_3nm, "../../shares/web/pressures/htdocs/data", "regions_offshore3nm_mol", driver="ESRI Shapefile")


# sea ice
ice <- raster(file.path(dir_M, "git-annex/Global/NCEAS-Pressures-Summaries_frazier2013/ice_mask_resampled"))
writeRaster(ice, "../../shares/web/pressures/htdocs/data/ice_mask_resampled.tif", format="GTiff", progress="text")

## put files in a zip
setwd("../../shares/web/pressures/htdocs/data")
zip(zipfile = "spatial_ohi_supplement",
               files = c("ocean.tif",
                         "regionData.csv",
                         "regions_offshore3nm_mol.dbf",
                         "regions_offshore3nm_mol.prj",
                         "regions_offshore3nm_mol.shp",
                         "regions_offshore3nm_mol.shx",
                         "sp_mol_raster_1km.tif",
                         "ice_mask_resampled.tif"))
file.remove(c("ocean.tif",
                         "regionData.csv",
                         "regions_offshore3nm_mol.dbf",
                         "regions_offshore3nm_mol.prj",
                         "regions_offshore3nm_mol.shp",
                         "regions_offshore3nm_mol.shx",
                         "sp_mol_raster_1km.tif",
                         "ice_mask_resampled.tif"))

setwd("../../../../../frazier/ftp_data")



```

**Figure 1. OHI regions**

For the Mollweide (A) and Latitude/Longitude (B) coordinate reference systems.

```{r region figs, echo=FALSE, fig.width=10, fig.height=4}

mol <- rasterGrob(readPNG("figures/global_map_Index_2016_mol.png"), interpolate=TRUE, y=.7)
mol_txt <- textGrob("A", gp=gpar(fontsize=12), just="left", x=0.15, y=0)

gcs <- rasterGrob(readPNG("figures/regions_gcs.png"), interpolate=TRUE, y=.7)
gcs_txt <- textGrob("B", gp=gpar(fontsize=12), just="left", x=0.15, y=0)
# c <- textGrob("c", gp=gpar(fontsize=12), just="left", x=0.15, y=0)
# d <- textGrob("d", gp=gpar(fontsize=12), just="left", x=0.15, y=0)
grid.arrange(mol_txt, gcs_txt, mol, gcs, nrow=2, ncol=2, heights=c(.2, 4))

```


**Table 1. OHI region metadata**

The region shapefiles include the following fields.

Field     | Includes    | Description
------------ | ----------- | -----------------------------------
rgn_typ      |  eez (N = 239), eez-disputed (15), eez-inland (5), fao (15), land (229), land-disputed (8), land-noeez (40) | Only the "eez" and "fao" regions are included in the global assessments.  "fao" regions refer to the [FAO Major Fishing Areas](http://www.fao.org/fishery/area/search/en), excluding eez regions.  The "eez" regions are based on the [exclusive economic zone boundaries](http://www.marineregions.org) of 220 regions and the [CCAMLR areas](https://www.ccamlr.org/en/organisation/convention-area-%E2%80%93-technical-description) for 19 regions in Antarctica.
ant_type     | eez (N = 220), eez-ccamlr (19), eez-disputed (15), eez-inland (5), fao (15), land (220), land-ccamlr (9), land-disputed (8), land-noeez (40) | This is similar to the rgn_typ variable but designates the Antarctica regions ("eez-ccamlr")
rgn_id       | numeric region identifiers (1 - 337) | 255 = disputed regions; 260-277 = fao regions; 301-337 = landlocked regions; 213 = Antarctica CCAMLR regions
ant_id       | numeric region identifiers | Same as rgn_id, except that Antarctica CCAMLR regions are identified
rgn_nam      | e.g., Antarctica, Peru    | Name of region
rgn_key      | e.g., ATA, UMI            | Short identifying code for region
are_km2      | numeric (<1, 30604795)   | Area of region in km^2^



## OHI pressure data
Many of the OHI pressures are available at the raster scale, and these data can be downloaded from Table 2).  To obtain the pressure score for each OHI region we average the raster cells within each region.  The raster data can be extracted at different spatial scales, which may be useful for regional OHI assessments.  

Resolution: ~1km 
Coordinate reference system: Mollweide 
Coverage: Global
Cell values: Scaled from 0 (low) - 1 (high-pressure)

```{r pressure data, echo=FALSE, eval=FALSE}

#################################
## ocean acidification data
#################################

years <- 2011:2015

for(year in years){ # year= 2011
file.copy(file.path(dir_M, "git-annex/globalprep/prs_oa/v2016/output", paste0("oa_prs_layer_", year, ".tif")),
          file.path("../../shares/web/pressures/htdocs/data", paste0("oa_prs_layer_", year, ".tif")))
}


setwd("../../shares/web/pressures/htdocs/data")
zip(zipfile = "acid", files = c(paste0("oa_prs_layer_", years, ".tif")))
file.remove(c(paste0("oa_prs_layer_", years, ".tif")))
setwd("../../../../../frazier/ftp_data")

############################
## SST
############################

years <- c("2008-2012", "2007-2011", "2006-2010", "2005-2009", "2004-2008")

for(year in years){ # year= 2011
file.copy(file.path(dir_M, "git-annex/globalprep/prs_sst/v2016/output", paste0("sst_", year, "_1985-1989.tif")),
          file.path("../../shares/web/pressures/htdocs/data", paste0("sst_", year, "_1985-1989.tif")))
}

setwd("../../shares/web/pressures/htdocs/data") 
zip(zipfile = "sst", files = c(paste0("sst_", years, "_1985-1989.tif")))
file.remove(c(paste0("sst_", years, "_1985-1989.tif")))
setwd("../../../../../frazier/ftp_data")

###############################
## SLR
###############################

years <- c(2011:2015)

for(year in years){ # year= 2011
file.copy(file.path(dir_M, "git-annex/globalprep/prs_slr/v2016/output", paste0("slr_", year, ".tif")),
          file.path("../../shares/web/pressures/htdocs/data", paste0("slr_", year, ".tif")))
}

setwd("../../shares/web/pressures/htdocs/data") 
zip(zipfile = "slr", files = c(paste0("slr_", years, ".tif")))
file.remove(c(paste0("slr_", years, ".tif")))
setwd("../../../../../frazier/ftp_data")

###############################
## UV
###############################

years <- c("2007_2011", "2008_2012", "2009_2013", "2010_2014", "2011_2015")

for(year in years){ # year= 2011
file.copy(file.path(dir_M, "git-annex/globalprep/prs_uv/v2016/output", paste0("uv_", year, "-2005_2009.tif")),
          file.path("../../shares/web/pressures/htdocs/data", paste0("uv_", year, "-2005_2009.tif")))
}

setwd("../../shares/web/pressures/htdocs/data") 
zip(zipfile = "uv", files = c(paste0("uv_", years, "-2005_2009.tif")))
file.remove(c(paste0("uv_", years, "-2005_2009.tif")))
setwd("../../../../../frazier/ftp_data")

###############################
## Trash
###############################

file.copy(file.path(dir_M, "git-annex/globalprep/cw_pressure_trash/v2015/output/weight_rescale.tif"),
          file.path("../../shares/web/pressures/htdocs/data/trash_weight_rescale.tif"))


###############################
## Artisanal fishing
###############################

years <- c("2003-2007", "2004-2008", "2005-2009", "2006-2010")

for(year in years){ # year= 2011
file.copy(file.path(dir_M, "git-annex/globalprep/prs_fish/v2016/out/artisanal", paste0("art_press_", year, ".tif")),
          file.path("../../shares/web/pressures/htdocs/data", paste0("art_press_", year, ".tif")))
}

setwd("../../shares/web/pressures/htdocs/data") 
zip(zipfile = "art_lb", files = c(paste0("art_press_", years, ".tif")))
file.remove(c(paste0("art_press_", years, ".tif")))
setwd("../../../../../frazier/ftp_data")

###############################
## Commercial fishing high bycatch
###############################

years <- c("2003-2007", "2004-2008", "2005-2009", "2006-2010")

for(year in years){ # year= 2011
file.copy(file.path(dir_M, "git-annex/globalprep/prs_fish/v2016/out/high_bycatch", paste0("hb_fish_pressure_", year, ".tif")),
          file.path("../../shares/web/pressures/htdocs/data", paste0("hb_fish_pressure_", year, ".tif")))
}

setwd("../../shares/web/pressures/htdocs/data") 
zip(zipfile = "comm_fis_hb", files = c(paste0("hb_fish_pressure_", years, ".tif")))
file.remove(c(paste0("hb_fish_pressure_", years, ".tif")))
setwd("../../../../../frazier/ftp_data")

###############################
## Commercial fishing low bycatch
###############################

years <- c("2003-2007", "2004-2008", "2005-2009", "2006-2010")

for(year in years){ # year= 2011
file.copy(file.path(dir_M, "git-annex/globalprep/prs_fish/v2016/out/low_bycatch", paste0("lb_fish_pressure_", year, ".tif")),
          file.path("../../shares/web/pressures/htdocs/data", paste0("lb_fish_pressure_", year, ".tif")))
}

setwd("../../shares/web/pressures/htdocs/data") 
zip(zipfile = "comm_fis_lb", files = c(paste0("lb_fish_pressure_", years, ".tif")))
file.remove(c(paste0("lb_fish_pressure_", years, ".tif")))
setwd("../../../../../frazier/ftp_data")

###############################
## Chemical pollution
###############################

years <- c(2009:2012)

# this one is a loner
file.copy(file.path(dir_M, "git-annex/globalprep/prs_chem/v2016/int/chemical_pollution_2013_scaled.tif"),
          file.path("../../shares/web/pressures/htdocs/data/chemical_pollution_2013_scaled.tif"))

for(year in years){ # year= 2011
file.copy(file.path(dir_M, "marine_threats/impact_layers_2013_redo/impact_layers/work/land_based/before_2007/raw_global_results/Frazier", paste0("chemical_pollution_", year, "_scaled.tif")),
          file.path("../../shares/web/pressures/htdocs/data", paste0("chemical_pollution_", year, "_scaled.tif")))
}

years <- 2009:2013
setwd("../../shares/web/pressures/htdocs/data") 
zip(zipfile = "chem_pollution", files = c(paste0("chemical_pollution_", years, "_scaled.tif")))
file.remove(c(paste0("chemical_pollution_", years, "_scaled.tif")))
setwd("../../../../../frazier/ftp_data")

###############################
## Nutrient pollution
###############################

years <- c(2009:2012)

# this one is a loner
file.copy(file.path(dir_M, "git-annex/globalprep/prs_land-based_nutrient/v2016/int/global_plumes_fert_2013_log_scaled_extend.tif"),
          file.path("../../shares/web/pressures/htdocs/data/global_plumes_fert_2013_log_scaled_extend.tif"))

for(year in years){ # year= 2011
file.copy(file.path(dir_M, "marine_threats/impact_layers_2013_redo/impact_layers/work/land_based/before_2007/raw_global_results/Frazier", paste0("global_plumes_fert_", year, "_raw_log_extend_scaled.tif")),
          file.path("../../shares/web/pressures/htdocs/data", 
                    paste0("global_plumes_fert_", year, "_raw_log_extend_scaled.tif")))
}

setwd("../../shares/web/pressures/htdocs/data") 
zip(zipfile = "nutrient_pollution", files = c("global_plumes_fert_2013_log_scaled_extend.tif", paste0("global_plumes_fert_", years, "_raw_log_extend_scaled.tif")))
file.remove(paste0("global_plumes_fert_", years, "_raw_log_extend_scaled.tif"))
file.remove("global_plumes_fert_2013_log_scaled_extend.tif")
setwd("../../../../../frazier/ftp_data")


```

**Table 2. OHI pressure raster data**

Links to download the raster data for several OHI pressures.  The "name" field links to a longer description of the data. The "description" field includes links to data preparation files (with R code).  NOTE: these large files can take a while to download and require a lot of memory!    

```{r pressure datatable, echo=FALSE, warning=FALSE}

link <- "http://ohi-science.org/ohi-global/layers.html#"

pressures <- c("cc_acid", "cc_sst", "cc_slr", "cc_uv", "po_trash", "fp_art_lb",
                      "fp_com_hb", "fp_com_lb", "po_chemicals", "po_nutrients")

p_data <- read.csv("../ohi-global/eez_layers_meta_data/layers_eez_base.csv") %>%
  filter(layer %in% pressures) %>%
  mutate(p_link = paste0(link, layer)) %>%
  mutate(name = paste0("[", name, "](", p_link, ")"))

p_prep <- read.csv("../ohi-global/eez_layers_meta_data/layers_eez_methods.csv") %>%
  filter(layer %in% pressures) %>%
  filter(variable == "dataprep_url_2016") %>%
  group_by(layer) %>%
  mutate(number = 1) %>%
  mutate(number = cumsum(number)) %>%
  mutate(number = paste0("data prep ", number)) %>%
  mutate(number = gsub(" 1", "", number)) %>%
  mutate(method = paste0("[", number, "](", variable_data, ")")) %>%
  summarize(method = paste(method, collapse = ", "))

download_data <- data.frame(layer = c("cc_acid", "cc_sst", "cc_slr", "cc_uv", "po_trash", "fp_art_lb", 
                                      "fp_com_hb", "fp_com_lb", "po_chemicals", "po_nutrients"),
                            Download = c("[5 years: 2011-2015, 1.7GB](https://mazu.nceas.ucsb.edu/pressures/data/acid.zip)",
                                         "[5 time periods: 2004-2008 to 2008-2012, 9.3GB](https://mazu.nceas.ucsb.edu/pressures/data/sst.zip)",
                                         "[5 years: 2011-2015, 99MB](https://mazu.nceas.ucsb.edu/pressures/data/slr.zip)",
                                         "[5 time periods: 2007-2011 to 2011-2015,  1.5GB](https://mazu.nceas.ucsb.edu/pressures/data/uv.zip)",
                                         "[1 year: 2015, 1.2GB](https://mazu.nceas.ucsb.edu/pressures/data/trash_weight_rescale.tif)",
                                         "[4 time periods: 2003-2007 to 2006-2010,  1.5GB](https://mazu.nceas.ucsb.edu/pressures/data/art_lb.zip)",
                                         "[4 time periods: 2003-2007 to 2006-2010,  2.6GB](https://mazu.nceas.ucsb.edu/pressures/data/art_lb.zip)",
                                         "[4 time periods: 2003-2007 to 2006-2010,  2.7GB](https://mazu.nceas.ucsb.edu/pressures/data/art_lb.zip)",
                                         "[5 years: 2009-2013, 2.3GB](https://mazu.nceas.ucsb.edu/pressures/data/chem_pollution.zip)",
                                         "[5 years: 2009-2013, 566MB](https://mazu.nceas.ucsb.edu/pressures/data/nutrient_pollution.zip)"))

p_data <- p_data %>%
  left_join(p_prep, by="layer") %>%
  mutate(description = paste0(description, " (", method, ")")) %>%
  left_join(download_data, by="layer") %>% 
  dplyr::select(Name=name, Description=description, Download)

kable(p_data)

```



