## OHI pressure data
Some of the processed OHI pressure data are available at the raster scale, which can be downloaded from Table 2.  To obtain the pressure score for each OHI region we average the raster cells within each region.  The raster data can be extracted at different spatial scales, which may be useful for regional OHI assessments.  

**An [example R script](https://rawgit.com/OHI-Science/ftp_data/master/tutorial/extract_spatial_data.html) describes how to extract and summarize raster data for regions.**

*Resolution:* ~1km 

*Coordinate reference system:* Mollweide 

*Coverage:* Global

*Cell values:* Scaled from 0 (low) - 1 (high-pressure)

```{r pressure data, echo=FALSE, eval=FALSE}

#################################
## ocean acidification data
#################################

years <- 2011:2015

for(year in years){ # year= 2011
file.copy(file.path(dir_M, "git-annex/globalprep/prs_oa/v2016/output", paste0("oa_prs_layer_", year, ".tif")),
          file.path("../../shares/web/data/htdocs/data", paste0("oa_prs_layer_", year, ".tif")))
}


setwd("../../shares/web/data/htdocs/data")
zip(zipfile = "acid", files = c(paste0("oa_prs_layer_", years, ".tif")))
file.remove(c(paste0("oa_prs_layer_", years, ".tif")))
setwd("../../../../../frazier/ftp_data")

############################
## SST
############################

years <- c("2008-2012", "2007-2011", "2006-2010", "2005-2009", "2004-2008")

for(year in years){ # year= 2011
file.copy(file.path(dir_M, "git-annex/globalprep/prs_sst/v2016/output", paste0("sst_", year, "_1985-1989.tif")),
          file.path("../../shares/web/data/htdocs/data", paste0("sst_", year, "_1985-1989.tif")))
}

setwd("../../shares/web/data/htdocs/data") 
zip(zipfile = "sst", files = c(paste0("sst_", years, "_1985-1989.tif")))
file.remove(c(paste0("sst_", years, "_1985-1989.tif")))
setwd("../../../../../frazier/ftp_data")

###############################
## SLR
###############################

years <- c(2011:2015)

for(year in years){ # year= 2011
file.copy(file.path(dir_M, "git-annex/globalprep/prs_slr/v2016/output", paste0("slr_", year, ".tif")),
          file.path("../../shares/web/data/htdocs/data", paste0("slr_", year, ".tif")))
}

setwd("../../shares/web/data/htdocs/data") 
zip(zipfile = "slr", files = c(paste0("slr_", years, ".tif")))
file.remove(c(paste0("slr_", years, ".tif")))
setwd("../../../../../frazier/ftp_data")

###############################
## UV
###############################

years <- c("2007_2011", "2008_2012", "2009_2013", "2010_2014", "2011_2015")

for(year in years){ # year= 2011
file.copy(file.path(dir_M, "git-annex/globalprep/prs_uv/v2016/output", paste0("uv_", year, "-2005_2009.tif")),
          file.path("../../shares/web/data/htdocs/data", paste0("uv_", year, "-2005_2009.tif")))
}

setwd("../../shares/web/data/htdocs/data") 
zip(zipfile = "uv", files = c(paste0("uv_", years, "-2005_2009.tif")))
file.remove(c(paste0("uv_", years, "-2005_2009.tif")))
setwd("../../../../../frazier/ftp_data")

###############################
## Trash
###############################

file.copy(file.path(dir_M, "git-annex/globalprep/cw_pressure_trash/v2015/output/weight_rescale.tif"),
          file.path("../../shares/web/data/htdocs/data/trash_weight_rescale.tif"))


###############################
## Artisanal fishing
###############################

years <- c("2003-2007", "2004-2008", "2005-2009", "2006-2010")

for(year in years){ # year= 2011
file.copy(file.path(dir_M, "git-annex/globalprep/prs_fish/v2016/out/artisanal", paste0("art_press_", year, ".tif")),
          file.path("../../shares/web/data/htdocs/data", paste0("art_press_", year, ".tif")))
}

setwd("../../shares/web/data/htdocs/data") 
zip(zipfile = "art_lb", files = c(paste0("art_press_", years, ".tif")))
file.remove(c(paste0("art_press_", years, ".tif")))
setwd("../../../../../frazier/ftp_data")

###############################
## Commercial fishing high bycatch
###############################

years <- c("2003-2007", "2004-2008", "2005-2009", "2006-2010")

for(year in years){ # year= 2011
file.copy(file.path(dir_M, "git-annex/globalprep/prs_fish/v2016/out/high_bycatch", paste0("hb_fish_pressure_", year, ".tif")),
          file.path("../../shares/web/data/htdocs/data", paste0("hb_fish_pressure_", year, ".tif")))
}

setwd("../../shares/web/data/htdocs/data") 
zip(zipfile = "comm_fis_hb", files = c(paste0("hb_fish_pressure_", years, ".tif")))
file.remove(c(paste0("hb_fish_pressure_", years, ".tif")))
setwd("../../../../../frazier/ftp_data")

###############################
## Commercial fishing low bycatch
###############################

years <- c("2003-2007", "2004-2008", "2005-2009", "2006-2010")

for(year in years){ # year= 2011
file.copy(file.path(dir_M, "git-annex/globalprep/prs_fish/v2016/out/low_bycatch", paste0("lb_fish_pressure_", year, ".tif")),
          file.path("../../shares/web/data/htdocs/data", paste0("lb_fish_pressure_", year, ".tif")))
}

setwd("../../shares/web/data/htdocs/data") 
zip(zipfile = "comm_fis_lb", files = c(paste0("lb_fish_pressure_", years, ".tif")))
file.remove(c(paste0("lb_fish_pressure_", years, ".tif")))
setwd("../../../../../frazier/ftp_data")

###############################
## Chemical pollution
###############################

years <- c(2009:2012)

# this one is a loner
file.copy(file.path(dir_M, "git-annex/globalprep/prs_chem/v2016/int/chemical_pollution_2013_scaled.tif"),
          file.path("../../shares/web/data/htdocs/data/chemical_pollution_2013_scaled.tif"))

for(year in years){ # year= 2011
file.copy(file.path(dir_M, "marine_threats/impact_layers_2013_redo/impact_layers/work/land_based/before_2007/raw_global_results/Frazier", paste0("chemical_pollution_", year, "_scaled.tif")),
          file.path("../../shares/web/data/htdocs/data", paste0("chemical_pollution_", year, "_scaled.tif")))
}

years <- 2009:2013
setwd("../../shares/web/data/htdocs/data") 
zip(zipfile = "chem_pollution", files = c(paste0("chemical_pollution_", years, "_scaled.tif")))
file.remove(c(paste0("chemical_pollution_", years, "_scaled.tif")))
setwd("../../../../../frazier/ftp_data")

###############################
## Nutrient pollution
###############################

years <- c(2009:2012)

# this one is a loner
file.copy(file.path(dir_M, "git-annex/globalprep/prs_land-based_nutrient/v2016/int/global_plumes_fert_2013_log_scaled_extend.tif"),
          file.path("../../shares/web/data/htdocs/data/global_plumes_fert_2013_log_scaled_extend.tif"))

for(year in years){ # year= 2011
file.copy(file.path(dir_M, "marine_threats/impact_layers_2013_redo/impact_layers/work/land_based/before_2007/raw_global_results/Frazier", paste0("global_plumes_fert_", year, "_raw_log_extend_scaled.tif")),
          file.path("../../shares/web/data/htdocs/data", 
                    paste0("global_plumes_fert_", year, "_raw_log_extend_scaled.tif")))
}

setwd("../../shares/web/data/htdocs/data") 
zip(zipfile = "nutrient_pollution", files = c("global_plumes_fert_2013_log_scaled_extend.tif", paste0("global_plumes_fert_", years, "_raw_log_extend_scaled.tif")))
file.remove(paste0("global_plumes_fert_", years, "_raw_log_extend_scaled.tif"))
file.remove("global_plumes_fert_2013_log_scaled_extend.tif")
setwd("../../../../../frazier/ftp_data")


```

**Table 2. OHI pressure raster data**

Links to download processed raster data for several OHI pressures.  Explore the links included in the table to learn more about the original data sources and how the data were prepared.  The "name" field links to a longer description of the data. The "description" field includes links to data preparation files (with R code).  NOTE: these large files can take a while to download and require a lot of memory!    

```{r pressure datatable, echo=FALSE, warning=FALSE}

link <- "http://ohi-science.org/ohi-global/layers.html#"

pressures <- c("cc_acid", "cc_sst", "cc_slr", "cc_uv", "po_trash", "fp_art_lb",
                      "fp_com_hb", "fp_com_lb", "po_chemicals", "po_nutrients")

p_data <- read.csv("../ohi-global/eez_layers_meta_data/layers_eez_base.csv") %>%
  filter(layer %in% pressures) %>%
  mutate(p_link = paste0(link, layer)) %>%
  mutate(name = paste0("[", name, "](", p_link, ")"))

p_prep <- read.csv("../ohi-global/eez_layers_meta_data/layers_eez_methods.csv") %>%
  filter(layer %in% pressures) %>%
  filter(variable == "dataprep_url_2016") %>%
  group_by(layer) %>%
  mutate(number = 1) %>%
  mutate(number = cumsum(number)) %>%
  mutate(number = paste0("data prep ", number)) %>%
  mutate(number = gsub(" 1", "", number)) %>%
  mutate(method = paste0("[", number, "](", variable_data, ")")) %>%
  summarize(method = paste(method, collapse = ", "))

download_data <- data.frame(layer = c("cc_acid", "cc_sst", "cc_slr", "cc_uv", "po_trash", "fp_art_lb", 
                                      "fp_com_hb", "fp_com_lb", "po_chemicals", "po_nutrients"),
                            Download = c("[5 years: 2011-2015, 1.7GB](https://ohi.nceas.ucsb.edu/data/data/acid.zip)",
                                         "[5 time periods: 2004-2008 to 2008-2012, 9.3GB](https://ohi.nceas.ucsb.edu/data/data/sst.zip)",
                                         "[5 years: 2011-2015, 99MB](https://ohi.nceas.ucsb.edu/data/data/slr.zip)",
                                         "[5 time periods: 2007-2011 to 2011-2015,  1.5GB](https://ohi.nceas.ucsb.edu/data/data/uv.zip)",
                                         "[1 year: 2015, 1.2GB](https://ohi.nceas.ucsb.edu/data/data/trash_weight_rescale.tif)",
                                         "[4 time periods: 2003-2007 to 2006-2010,  1.5GB](https://ohi.nceas.ucsb.edu/data/data/art_lb.zip)",
                                         "[4 time periods: 2003-2007 to 2006-2010,  2.6GB](https://ohi.nceas.ucsb.edu/data/data/art_lb.zip)",
                                         "[4 time periods: 2003-2007 to 2006-2010,  2.7GB](https://ohi.nceas.ucsb.edu/data/data/art_lb.zip)",
                                         "[5 years: 2009-2013, 2.3GB](https://ohi.nceas.ucsb.edu/data/data/chem_pollution.zip)",
                                         "[5 years: 2009-2013, 566MB](https://ohi.nceas.ucsb.edu/data/data/nutrient_pollution.zip)"))

p_data <- p_data %>%
  left_join(p_prep, by="layer") %>%
  mutate(description = paste0(description, " (", method, ")")) %>%
  left_join(download_data, by="layer") %>% 
  dplyr::select(Name=name, Description=description, Download)

kable(p_data)

```
